A Switching Planner for Combined Task and Observation Planning

### 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
typo

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(:init)$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the unique abstract state with probability $1.0$ -- E.g., for
$\pp{DORA}$ $s_0 \equiv \{(= (\pp{is-in}~\pp{DORA})~\pp{kitchen})\}$.

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
Executing $\assumptiveS{i}$ in $s$ effects a transition to a
successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

Because $\state_0$ is abstract, and because actions can have
preconditions that are not uniformly true in the underlying
(resp. abstract) belief-state, we must make two further adjustments to
the problem posed to the sequential planner. First, if
$\action\in\actions$ has a precondition that is not true with
probability $1$, then before it is executed blindly, the sequential
continual planning switches to a decision-theoretic session, discussed in
detail below. Overall, by scheduling $\action$ the sequential planner
makes a naive assumption about the possibility of the agent eventually
``committing'' to the truth values of specific state
propositions. Thus, the role of decision-theoretic plans is to verify
the validity of assumptions about the runtime state in sequential
planning. Second, if the truth value of $\prop$ is not known, and no
assumption is made about it in the current abstract state, then if
$\prop \in
\poss(\action)$ we have that action $\action$ is not applicable.

%% Because $\state_0$ is abstract, and because actions can have
%% preconditions that are not uniformly true in the underlying
%% (resp. abstract) belief-state, we must make two further adjustments to
%% the problem posed to the sequential planner. First, if
%% $\action\in\actions$ has a precondition that is not true with
%% probability $1$, then before it is executed blindly, the sequential
%% continual planning switches to a decision-theoretic mode, discussed in
%% detail below. Overall, by scheduling $\action$ the sequential planner
%% makes a naive assumption about the possibility of the agent eventually
%% committing ---for the purpose of the current task--- to the truth
%% values of specific state propositions. Second, if the truth value of
%% $\prop$ is not known, and no assumption is made about it in the
%% current abstract state, then if $\prop \in \poss(\action)$ we have
%% that action $\action$ is not applicable.

%% In other words, the value of a proposition in an abstract
%% state is either {\em true}, {\em false}, or {\em unspecified} (i.e.,
%% \#). In the last case actions whose execution is preconditioned on the
%% term being truth functional are not available.

@@

In goal directed problems, where the only non-zero reward is received
at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
criteria corresponds exactly to that of the infamous degenerate
hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

@@

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ block.  This is constructed by first
constructing a tree which only includes terms from the {\em relaxed}
visitation characterised by the {\em assumptive} actions in the
current sequential plan. For example, if $\pp{DORA}$'s plan is:

@@

In a second step, we iteratively refine the {\em relaxed} block by
adding terms from the original block while the resulting state space
remains of a practicable size. Of course, it is expensive to measure
size exactly, thus in practice we estimate it to be
$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
propositions in the abstraction induced by the proposed tree. Terms
are proposed for addition because they are probabilistic, and because
they were omitted in the {\em relaxed} visitation characterised by
{\em assumptive} actions in the sequential plan. Each iteration, we
add one candidate term with the largest number of parent terms already
in the candidate tree.

>>>>

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the unique abstract state with probability $1.0$ -- E.g., for
$\pp{DORA}$:

\[
\begin{array}{l}
s_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
\end{array}
\]

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $p_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. The physics of these actions is as
follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
Executing $\assumptiveS{i}$ in $s$ effects a transition to a successor
state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with probability
$1 - p_i$. Here, $s^{T_i}$ is the union of $s$ with atomic terms from
$T_i$ and conflicting propositions from $s$ removed. For example, if
the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
delete $(=(\pp{is-in}~\pp{box})~\#)$. State $s^\bot$ is a sink, where
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

Because $\state_0$ is abstract, and because actions can have
preconditions that are not uniformly true in the underlying
(resp. abstract) belief-state, we must make two further adjustments to
the problem posed to the sequential planner. First, if
$\action\in\actions$ has a precondition that is not true with
probability $1$ in $\bstate_0$, then before it is executed blindly,
the sequential continual planning switches to a decision-theoretic
session, discussed in detail below. Overall, by scheduling $\action$
the sequential planner makes a naive assumption about the possibility
of the agent eventually ``committing'' to the truth values of specific
state propositions in $\poss(\action)$. The role of decision-theoretic
plans is to examine the validity of assumptions about the runtime
state of sequential planning. Second, if the truth value of $\prop$ is
not known, and no assumption is made about it in the current abstract
state, then if $\prop \in
\poss(\action)$ we have that action $\action$ is not applicable.

@@

In goal directed problems, where the only non-zero reward is received
at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
criteria corresponds exactly to that of
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

@@

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ declaration.  We first construct a
tree which only includes terms from the {\em relaxed} visitation
characterised by the {\em assumptive} actions in the current
sequential plan. For example, if $\pp{DORA}$'s plan is:

@@

In a second step, we iteratively refine the above{\em relaxed}
declaration by adding terms from the original statement of
$(\pp{:init})$ while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
and in practice we estimate this to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction that
were not assigned the {\em unknown} value \#. Terms are proposed for
addition because they are probabilistic, and because they were omitted
in the {\em relaxed} visitation characterised by {\em assumptive}
actions in the sequential plan. Each iteration, we add one candidate
term with the largest number of parent terms already in the candidate
tree.
